---
title: "Activity03"
author: "Pep Canyadell"
date: "8/9/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

# Install and load necessary packages
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}

if (!requireNamespace("tidymodels", quietly = TRUE)) {
  install.packages("tidymodels")
}

if (!requireNamespace("GGally", quietly = TRUE)) {
  install.packages("GGally")
}

library(tidyverse)
library(tidymodels)
library(GGally)



```


```{r}

# Load the mtcars dataset
data("mtcars")

# Assign the dataset to a data frame
df <- mtcars

```


```{r}

# Create pairwise relationship plots using GGally
df %>%
  select(mpg, hp, wt) %>%
  ggpairs()



```

```{r}

# Fit a multiple linear regression model
lm_spec <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm")

mlr_mod <- lm_spec %>% 
  fit(mpg ~ hp + wt, data = df)

# Display model parameters
tidy(mlr_mod)


```

```{r}

# Create a 3-D scatterplot using ggplot2
library(ggplot2)

ggplot(df, aes(x = hp, y = wt, z = mpg)) +
  geom_point() +
  labs(x = "Horsepower", y = "Weight", z = "Miles Per Gallon") +
  theme_minimal()


```


# Procedure:

Task 1: Opened the RMarkdown document:

I began by opening a special document in RStudio called "activity03.Rmd."

Task 2: Loaded the necessary packages:

I brought in some extra tools to help me work. It was like getting a toolkit ready. I checked if the tools were already there, and if not, I quickly installed and used them.

Task 3: Loaded the data:

I picked a collection of information about cars, like miles per gallon, horsepower, and weight. I named this collection "df."

Task 4: Created pairwise relationships:

I made pictures that showed how these car things were related to each other. For example, I showed how miles per gallon and horsepower were connected.

Task 5: Fitted the multiple linear regression model:

I used a special math trick to figure out if one thing could help predict another thing. It was like guessing if knowing the horsepower and weight of a car could help guess its miles per gallon.

Challenge: Created 3-D plots:

I drew a fancy picture with three axes. Imagine I had toy cars with numbers for how much horsepower, weight, and miles per gallon they had. I put the cars in this picture where each axis showed one of those numbers.
